---
title: "Lab 2 -- Beta-Binomial Distribution"
author: "Rebecca C. Steorts"
date: "January 2018"
output: pdf_document
---

In class, you saw the Binomial-Beta model. We will now use this to solve a very real problem! Suppose I wish to determine whether the probability that a worker will fake an illness is truly 1\%. Your task is to assist me! Tasks 1--3 will be completed in lab and tasks 3--5 should be completed in your weekly homework assignment. You should still upload task 3 even though this will be worked through in lab! 

# Task 1

Let's start by quickly deriving the Beta-Binomial distribution. 

We assume that 
$$X\mid \theta \sim \text{Binomial} (\thseta)$$,
$$\theta \sim \text{Beta}(a,b),$$
where $a,b$ are assumed to be known parameters. What is the posterior distribution of $\theta \mid X$?

\begin{align}
p(\theta \mid X) &\propto 
p(X \mid \theta) p(\theta) \\
&\propto \theta^{x} 
(1 - \theta)^{(n-x)} \times \theta^{(a-1)} (1 - \theta)^{(b-1)}\\
&\propto \theta^{x + a -1} (1 - \theta)^{(n-x + b -1)}.
\end{align}

This implies that 
$$\theta \mid X \sim \text{Beta}(x+a,n-x+b).$$



# Task 2

Simulate some data using the \textsf{rbinom} function of size $n = 100$ and probability equal to 1\%. Remember to \textsf{set.seed(123)} so that you can replicate your results.

The data can be simulated as follows:
```{r,echo=TRUE}
# set a seed
set.seed(123)
# create the observed data
obs.data <- rbinom(n = 100, size = 1, prob = 0.01)
# inspect the observed data
head(obs.data)
tail(obs.data)
length(obs.data)
```

# Task 3

Write a function that takes as its inputs that data you simulated (or any data of the same type) and a sequence of $\theta$ values of length 1000 and produces Likelihood values based on the Binomial Likelihood. Plot your sequence and its corresponding Likelihood function.

The likelihood function is given below. Since this is a probability and is only valid over the interval from $[0, 1]$ we generate a sequence over that interval of length 1000.

You have a rough sketch of what you should do for this part of the assignment. Try this out in lab on your own. 

```{r, echo = TRUE}
### Bernoulli LH Function ###
# Input: obs.data, theta
# Output: bernoulli likelihood
LH.func = function(obs.data,theta){
  x = sum(obs.data)
  N = length(obs.data)
  LH = (theta^x)*(1-theta)^{N-x}
  return(LH)
}


### Plot LH for a grid of theta values ###
# Create the grid #
theta.sim = seq(0,1,length.out=1000)
# Store the LH values
LH.sim = LH.func(obs.data,theta.sim)
# Create the Plot
plot(x=theta.sim,y=LH.sim,type='l',main='Likelihood Profile',xlab="Simulated Support of Theta",ylab='Likelihood')
```



# Task 4 (To be completed for homework)

Write a function that takes as its inputs  prior parameters \textsf{a} and \textsf{b} for the Beta-Bernoulli model and the observed data, and produces the posterior parameters you need for the model. \textbf{Generate and print} the posterior parameters for a non-informative prior i.e. \textsf{(a,b) = (1,1)} and for an informative case \textsf{(a,b) = (3,1)}}.

```{r}
#Function for posterior parameters 
PostPram.func = function(a,b,obs.data){
  x = sum(obs.data)
  N = length(obs.data)
  Post.a = x + a
  Post.b = N - x + b
  ret = list("a" = Post.a, "b" = Post.b)
  return(ret)
}

##Generate and Print Posterior Parameters
#Case 1: non-informative prior 
Pram.NonInf = PostPram.func(1,1,obs.data)
Pram.NonInf
```
```{r}
#Case 2: informative prior 
Pram.Info = PostPram.func(3,1,obs.data)
Pram.Info
```



# Task 5 (To be completed for homework)

Create two plots, one for the informative and one for the non-informative case to show the posterior distribution and superimpose the prior distributions on each along with the likelihood. What do you see? Remember to turn the y-axis ticks off since superimposing may make the scale non-sense.


```{r}
#create support
p = seq(0,1, length=1000)

##plot Non-Informative Prior and Posterior along with likelihood
#plot posterior
plot(p, dbeta(p, Pram.NonInf$a,Pram.NonInf$b),lwd=4, ylab='density', type ='l', col='yellow', main = "Non-Informative Prior", xlab = "theta", yaxt='n')
par(new=TRUE)
#plot prior
plot(p, dbeta(p, 1, 1), col='purple', type = 'l',ylab = "",yaxt='n',xlab="") 
par(new=TRUE)
#plot likelihood
plot(p, LH.sim, col='red', type = 'l', ylab = "",yaxt='n',xlab="") 


#add legend
legend("topright", c( "Posterior","Prior",'Likelihood'),lty=c(1,1,1),col=c("yellow","purple","red"))

```

```{r}
#create support
p = seq(0,1, length=1000)

#plot Informative Prior and Posterior along with likelihood
plot(p, dbeta(p, Pram.Info$a,Pram.Info$b),lwd=4, ylab='density', type ='l', col='yellow', main = "Informative Prior", xlab = "theta", yaxt='n')
par(new=TRUE)
#plot prior
plot(p, dbeta(p, 3, 1), col='purple', type = 'l',ylab = "",yaxt='n',xlab="") 
par(new=TRUE)
#plot likelihood
plot(p, LH.sim, col='red', type = 'l', ylab = "",yaxt='n',xlab="") 


#add legend
legend("topright", c( "Posterior","Prior",'Likelihood'),lty=c(1,1,1),col=c("yellow","purple","red"))
```

